---
title: "Figure 5"
author: "Robert Froemel & Lars Velten"
date: "2025-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data from figshare

```{r load_data}
load(url("https://figshare.com/ndownloader/files/53009321"))
load("additional_data/Figure5_LibHAnnotation.rda") #also load additional annotation / design info for Lib H
combined <- readRDS("Figure2_TF_data.rds") #and the summaries generated in figure 2

```

```{r load_packages, warning=F, message=F}
require(ggplot2)
require(ggrepel)
require(plyr)
require(reshape2)
require(gridExtra)
require(stringr)
require(motifmatchr)
require(TFBSTools)
require(SummarizedExperiment)
```

## Figure 5B

Simply determine % of sequences active above background for each factor combination and cell state, for LibB and LibH.

```{r fig5b, fig.width=2.2,fig.height=2.35}

heptad <- unique(c(mpra.data$HSPC.libB$DATA$TF1.name,unique(mpra.data$HSPC.libB$DATA$TF2.name)))

MULTI <- merge(mpra.data$HSPC.libH$DATA, libH.meta[, c("CRS", "Design")])

all_bad_sites <- subset(mpra.data$HSPC.libH$CONTROLS.GENERAL,ControlType == "ScrambleTFBS")
bad_range <- ddply(as.data.frame(all_bad_sites), "clusterID",summarise, q05.background = quantile(mean.norm.adj, 0.05), q95.background = quantile(mean.norm.adj,0.95))
rownames(bad_range) <- bad_range$clusterID

MULTI <- merge(MULTI, bad_range)
MULTI$active <- with(MULTI, mean.norm.adj > q95.background)
MULTI$repressed <- with(MULTI, mean.norm.adj < q05.background)

TF.count <- sapply(pwm.names, function(x) str_count(MULTI$Design, sprintf("%s_",x)))

MULTI <- cbind(MULTI, TF.count)


MULTI$Fli1Spi1 <- with(MULTI, paste(Fli1 > 0, Spi1 > 0))
perc.active <- ddply(MULTI, c("Fli1Spi1","clusterID"), summarise, perc.act = mean(active), perc.repr = mean(repressed), N = length(active))


pairs <- do.call(rbind, lapply(names(cellstate.map), function(ct) {
  ct.m <- MULTI$clusterID == ct
  do.call(rbind, lapply(heptad, function(tf1) {
    tf1.m <- MULTI[, tf1] > 0
    do.call(rbind, lapply(heptad, function(tf2) {
      tf2.m <- MULTI[, tf2] > 0
      relevant <- MULTI[ct.m & tf1.m & tf2.m,  ]
      data.frame(clusterID = ct, TF1 = tf1, TF2 = tf2, perc.active = mean(relevant$active), perc.repr = mean(relevant$repressed), N = nrow(relevant))
    }))
  }))
}))
pairs$tfcombo <- with(pairs, sprintf("%s-%s", TF1, TF2))

multi.vs.libH <- merge(pairs, combined)
cellstate.colors["State_7M"] <- "#AF7520"

ggplot(aes(x = 100*n_active, y = 100*perc.active), data = subset(multi.vs.libH,N>10 )) + 
  geom_point( color = "grey60", size=0.75) + 
  geom_point(aes(color = clusterID), data = subset(multi.vs.libH, TF2 == "Fli1" & TF1 == "Spi1")) + 
  ylab("% active, Pair (Lib B)") +
  scale_x_log10()+ scale_y_log10()+
  scale_color_manual(values = cellstate.colors, labels = cellstate.map, name = "Example:\nFli1-Spi1 in...") + 
  scale_shape_discrete(labels = cellstate.map) + 
  xlab("% active\nComplex Synthetic (Lib H)")  + theme_bw() + theme(panel.grid = element_blank(), legend.position = "none") + 
  scale_size_continuous(range= c(0.5,1.5), breaks = round(seq(min(multi.vs.libH$N), max(multi.vs.libH$N), length.out = 4  )))
```

## Figure 5C

Simply determine % of sequences active above background for each factor combination and cell state, for LibB and LibG. In LibG, factor combinations are identified by motifmatchr.

```{r fig5c, fig.width=3,fig.height=2.35}
N <- mpra.data$HSPC.libG$DATA

motif_scores <- matchMotifs(motifs, N$Seq,  genome = BSgenome.Hsapiens.UCSC.hg38,out = "scores")
N  <- cbind(N, as.matrix(assay(motif_scores)))
N$Fli1Spi1 <- with(N, paste(Fli1 > 0, Spi1 > 0))


all_bad_sites <- subset(mpra.data$HSPC.libG$CONTROLS.GENERAL,ControlType == "ScrambleTFBS")
bad_range <- ddply(as.data.frame(all_bad_sites), "clusterID",summarise, q05.background = quantile(mean.norm.adj, 0.05), q95.background = quantile(mean.norm.adj,0.95))
rownames(bad_range) <- bad_range$clusterID

N <- merge(N, bad_range)
N$active <- with(N, mean.norm.adj > q95.background)
N$repressed <- with(N, mean.norm.adj < q05.background)

perc.active <- ddply(N, c("Fli1Spi1","clusterID"), summarise, perc.act = mean(active), perc.repr = mean(repressed), N = length(active))


#do this systematically for all pairs to gain correlations with libH
pairs <- do.call(rbind, lapply(names(cellstate.map), function(ct) {
  ct.m <- N$clusterID == ct
  do.call(rbind, lapply(heptad, function(tf1) {
    tf1.m <- N[, tf1] > 0
    do.call(rbind, lapply(heptad, function(tf2) {
      tf2.m <- N[, tf2] > 0
      relevant <- N[ct.m & tf1.m & tf2.m,  ]
    data.frame(clusterID = ct, TF1 = tf1, TF2 = tf2, perc.active = mean(relevant$active), perc.repr = mean(relevant$repressed), N = nrow(relevant))
  }))
}))
}))
pairs$tfcombo <- with(pairs, sprintf("%s-%s", TF1, TF2))

nat.vs.libH <- merge(pairs, combined)


ggplot(aes(x = 100*n_active, y = 100*perc.active), data = subset(nat.vs.libH,N>10 )) + 
  geom_point( aes( size = N), color = "grey60") + 
  geom_point(aes(size = N, color = clusterID), data = subset(nat.vs.libH, TF2 == "Fli1" & TF1 == "Spi1")) + 
  ylab("% active, Pair (Lib B)") +
  scale_x_log10()+ scale_y_log10()+
  scale_color_manual(values = cellstate.colors, labels = cellstate.map, name = "Example:\nFli1-Spi1 in...") + 
  scale_shape_discrete(labels = cellstate.map) + 
  xlab("% active\nGenome (Lib G)")  + theme_bw() + theme(panel.grid = element_blank()) + 
  scale_size_continuous(name = "N obs.\n(Lib G)",range= c(0.2,3.5), breaks = round(seq(10, max(nat.vs.libH$N), length.out = 4  )))

```

## Figure 5D

```{r fig5d, fig.width = 2.2,fig.width=4}
    active_by_celltype <- ddply(as.data.frame(N), c("Fli1Spi1","clusterID"), summarise,
                                n_active = mean(mean.norm.adj > bad_range[unique(clusterID),"q95.background"]),
                                n_repressed = mean(mean.norm.adj < bad_range[unique(clusterID),"q05.background"]))

qplot(x = Fli1Spi1, y = clusterID, fill = 100*n_repressed, data = active_by_celltype, geom="tile") + 
  scale_x_discrete(labels = c("FALSE FALSE" = "Neither motif", "FALSE TRUE" = "Fli1, no Spi1", "TRUE FALSE" = "Spi1, no Fli1", "TRUE TRUE" = "Both motifs")) +
  scale_y_discrete(labels = cellstate.map) + 
  theme_bw() + scale_fill_gradientn(name = "% sequences\nrepressed", colours = c("black","blue", "green")) +
  theme(axis.text.x = element_text(angle=45, hjust=1), axis.title =  element_blank())  

```

## Figure 5F

Code for training the deep learning model is on https://github.com/veltenlab/MPRA_prediction . During training, the predictions on the held out testsets are written to a csv file, which we here just read in, and merge with the measured values to compute the test set R2.

```{r fig5f,fig.width=2,fig.height=2}

rbind_overlaps <- function(...) {
  entries <- list(...)
  rbind_o_internal <- function(a,b){
    usecols <- intersect(colnames(a),colnames(b))
    rbind(a[,usecols],b[,usecols])
  }
  out <- rbind_o_internal(entries[[1]],entries[[2]])
  if (length(entries) > 2) {
    for (i in 3:length(entries)) {
      out <-rbind_o_internal(out,entries[[i]])
    }
  }
  return(out)
}

predictions <- readRDS("additional_data/testset_predictions.rds")
all_data <- rbind_overlaps(
  mpra.data$HSPC.libA$DATA,
  mpra.data$HSPC.libB$DATA,
  mpra.data$HSPC.libC$DATA
)

all_data$Seq <- toupper(all_data$Seq)

all_data <-  merge(all_data, predictions, all.x = T)

R2 <- ddply(all_data, c("Library"), summarise, 
            R2_batch = cor(deep.batch, mean.norm.adj, use = "pairwise.complete.obs")^2, 
            R2_tech = cor(norm.1.adj, norm.2.adj, use = "pairwise.complete.obs")^2)
liblabel <- c("LibA"="Lib A", "LibB" = "Lib C", "LibH" ="Lib B")  #the labels used internally are different from the ones in the paper
R2$label = liblabel[as.character(R2$Library)]

 a <- ggplot(aes(x= label, y = R2_batch),data=R2 ) + 
  geom_col() + ylab("R2 (10-fold CV)") + xlab("Library") + theme_bw() + theme(panel.grid = element_blank()) + 
  geom_point(aes(y = R2_tech), shape =4)
 print(a)
```

```{r fig5fpdf}
pdf("panels/figure5f.pdf",width=2, height=2)
print(a)
dev.off()
```

## Figure 5G

After the model is trained, we can use the functions from https://github.com/veltenlab/MPRA_prediction/blob/master/query_model_client.R to get predictions on Lib G and Lib H.

```{r fig5g, fig.width=3.6, fig.height=2.4}

#The code for queryign the model would be
# libG <- predict_deep_model_batch(unique(mpra.data$HSPC.libG$DATA$Seq), "LibG", server = "127.0.0.1"))

#here we just load the predictions
libG.prediction <-as.data.frame(readRDS("additional_data/libG_predictions.rds"))
libG.prediction$Seq <- toupper(rownames(libG.prediction))
mpra.data$HSPC.libG$DATA$Seq <- toupper(mpra.data$HSPC.libG$DATA$Seq)
libG.prediction <- melt(libG.prediction, id.vars = c( "Seq"), value.name = "deep.batch", variable.name = "clusterID")

holdout <- merge(mpra.data$HSPC.libG$DATA, libG.prediction)

#add info on motifs
motif_scores <- matchMotifs(motifs, holdout$Seq,  genome = BSgenome.Hsapiens.UCSC.hg38,out = "scores")
holdout  <- cbind(holdout, as.matrix(assay(motif_scores)))

holdout$nmotif <- apply(assay(motif_scores) > 0, 1,sum)

#same thing for LibH
libH.prediction <- readRDS("additional_data/libH_predictions.rds")
mpra.data$HSPC.libH$DATA$Seq <- substr(toupper(mpra.data$HSPC.libH$DATA$Seq),5,250)

holdoutH <- merge(mpra.data$HSPC.libH$DATA, libH.prediction)
holdout.complete <- rbind_overlaps(holdout, holdoutH)
holdout.complete$Library <- ifelse(holdout.complete$Library == "LibB", "LibH", "LibG") #internally we used different codes for the libraries

R2_holdout <- ddply(holdout.complete, c( "Library"), summarise, 
                    R2_batch = cor(deep.batch, mean.norm.adj, use = "pairwise.complete.obs")^2, 
                    R2_tech = cor(norm.1.adj, norm.2.adj, use = "pairwise.complete.obs")^2, 
                    nobs = length(Seq))
R2_holdout_nmotif <- ddply(holdout.complete, c( "Library", "nmotif"), summarise, 
                    R2_batch = cor(deep.batch, mean.norm.adj, use = "pairwise.complete.obs")^2, 
                    R2_tech = cor(norm.1.adj, norm.2.adj, use = "pairwise.complete.obs")^2,
                    nobs = length(Seq))


ggplot(aes(x = as.factor(nmotif), y = R2_batch), data = R2_holdout_nmotif) + 
  coord_cartesian() + geom_point(aes(size=nobs)) +
  scale_y_sqrt(limits = c(0, max(R2_holdout_nmotif$R2_batch)), breaks = c(0,0.1,0.3,0.5))+
  scale_size_continuous(range =c(0.5,2.5), name = "N observations", breaks = round(seq(min(R2_holdout_nmotif$nobs), max(R2_holdout_nmotif$nobs), length.out=4))) +
  xlab("N motifs (42 TFs)") + ylab("R2") +
  theme_bw() + theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank(), legend.position = "bottom") + 
  geom_hline(aes(yintercept = R2_tech),data= R2_holdout, linetype = 2) + facet_wrap(~Library, scales= "free_x") 
```

```{r sinfo}
print(sessionInfo())
```